Data Cleaning

### 2025-12-09 Data Cleaning.Rmd
Started into Random Forest model for classification

df_lag_incdec <- read.csv("RWAR_dosing_lag_incdec.csv")

df_lag_incdec$FeCl_change <- as.factor(df_lag_incdec$FeCl_change)
df_lag_incdec$Polymer_change <- as.factor(df_lag_incdec$Polymer_change)

summary(df_lag_incdec)

df_lag_incdec <- df_lag_incdec |> 
  mutate(Timestamp = ymd(Timestamp))

df_lag_incdec |>
  select(-c(Timestamp)) |>
  group_by(FeCl_change) |>
  summarise(mean_Temp = mean(mean_Temp),
            mean_pH = mean(mean_pH),
            mean_cond = mean(mean_Conductivity),
            mean_turbidity = mean(mean_Turbidity))

@@

@@

```{r}
new_df <- df_lag_incdec |>
  rename(FeCl_dose = Ferric.chloride..mg.L,
         Polymer_dose = Cationic.Polimer..mg.L)

write.csv(new_df, "RWAR_dosing_lag_incdec.csv")

>>>>

# Pull in data
df_lag_incdec <- read.csv("RWAR_dosing_lag_incdec.csv") |> 
  mutate(Timestamp = ymd(Timestamp))
df_lag_incdec$FeCl_change <- as.factor(df_lag_incdec$FeCl_change)
df_lag_incdec$Polymer_change <- as.factor(df_lag_incdec$Polymer_change)

# Look at average water characteristics for each FeCl dose change group
df_lag_incdec |>
  select(-c(Timestamp)) |>
  group_by(FeCl_change) |>
  summarise(count = n(),
            mean_Temp = mean(mean_Temp),
            mean_pH = mean(mean_pH),
            mean_cond = mean(mean_Conductivity),
            mean_turbidity = mean(mean_Turbidity))

# Look at average water characteristics for each polymer dose change group
df_lag_incdec |>
  select(-c(Timestamp)) |>
  group_by(Polymer_change) |>
  summarise(count = n(),
            mean_Temp = mean(mean_Temp),
            mean_pH = mean(mean_pH),
            mean_cond = mean(mean_Conductivity),
            mean_turbidity = mean(mean_Turbidity))

@@

## TRAIN/TEST DATA
train_df <- df_lag_incdec |> filter(year(Timestamp) <= 2019)
test_df  <- df_lag_incdec |> filter(year(Timestamp) == 2020)

## RANDOM FOREST

rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
  set_engine("randomForest", importance = TRUE) |>
  set_mode("classification")

rf_fit <- rf_spec |>
  fit(high_sales ~ ., data = df_train)

vip(rf_fit) ## still Price and Shelf location are the most important

rf_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = high_sales, estimate = .pred_class)

# 5. Compare the OOB confusion matrix to your test confusion matrix. [**Hint:** The `confusion` element of the model output fit is OOB.]

## OOB:
rf_fit |>
  extract_fit_engine() %>%
  .$confusion

## test confusion
rf_fit |>
  augment(new_data = df_test) |>
  conf_mat(truth = high_sales, estimate = .pred_class)
```

@@
